Lensing in the Schwarzschild Spacetime Applicable to Neutron Stars

### 2023-09-29 src/tex/ms.tex
added comments to the geometry part

To convert this world-line tangent vector to the \emph{observer} coordinate
system, let the inclination of the neutron star spin be $\iota$ and the
obliquity (measered ccw from ``vertical'' in the observer frame) be $\omega$.
Then the neutron star spin three-vector has cartesian components in the observer
frame of 
\begin{equation}
    \vec{s} = \left( \cos \iota, -\sin \iota \sin \omega, \sin \iota \cos \omega \right).
\end{equation}
%%%
\begin{figure}
    \centering
        \includegraphics[width=0.9\columnwidth]{obs_cart_coord.png}
        \caption{The coordinates in the observer frame.}
        \label{fig:testsim}
    \end{figure}
%%%
If the incoming ray strikes the surface at a point with colatitude $\theta_0$
and longitude $\phi_0$, then the vector pointing to the impact point has
cartesian components in the observer frame of 
\begin{equation}
    \vec{r}_0 = \left( \cos \phi_0 \sin \theta_0, \sin\phi_0 \sin \theta_0, \cos \theta_0 \right).
\end{equation}
The instantaneous world-line tangent of the observer fixed to the NS surface at
the point of impact is given in cartesian coordinates by 
\begin{equation}
    \vec{t}_0 = \frac{2 \pi r_{NS}}{P} \frac{\partial \chi^t}{\partial \lambda} \vec{s} \times \vec{r}_0. 
\end{equation}
In the observer spherical coordinate system the components of the tangent vector
are given by 
\begin{eqnarray}
    \frac{\mathrm{d}\chi^t}{\mathrm{d} \lambda} & \simeq & 1 + \frac{1}{2}  \vRot^2 \sin^2 \theta_0 + \frac{1 + \frac{3}{2} \vRot^2 \sin^2 \theta_0}{\rNS} \\
    \frac{\mathrm{d} \chi^r}{\mathrm{d} \lambda} & = & 0 \\
    \frac{\mathrm{d} \chi^\theta}{\mathrm{d} \lambda} & = & \frac{\mathrm{d} \chi^t}{\mathrm{d}\lambda} \frac{2\pi}{P} \left( - \cos\iota \sin \phi_0 - \cos\phi_0 \sin \iota \sin \omega \right) \\
    \frac{\mathrm{d} \chi^\phi}{\mathrm{d} \lambda} & = & \frac{\mathrm{d} \chi^t}{\mathrm{d}\lambda} \frac{2\pi}{P} \left( -\cos\iota \cos \phi_0 \cot \theta_0  + \sin \iota \left( \cos \omega + \cot \theta_0 \sin \phi_0 \sin \omega \right) \right)
\end{eqnarray}

>>>>

To convert this world-line tangent vector to the \emph{observer} coordinate
system, let the inclination of the neutron star spin be $\iota$ and the
obliquity (measered ccw from ``vertical'' in the observer frame) be $\omega$.
Then the neutron star spin three-vector has cartesian components in the observer
frame of 
\begin{equation}
    \vec{s} = \left( \cos \iota, -\sin \iota \sin \omega, \sin \iota \cos \omega \right).
\end{equation}
See the Figure \ref{fig:obs_cart}.
%%%
\begin{figure}
    \centering
        \includegraphics[width=0.9\columnwidth]{obs_cart_coord.png}
        \caption{The coordinates in the observer frame.}
        \label{fig:obs_cart}
    \end{figure}
%%%
If the incoming ray strikes the surface at a point with colatitude $\theta_0$
and longitude $\phi_0$, then the vector pointing to the impact point has
cartesian components in the observer frame of 
\begin{equation}
    \vec{r}_0 = \left( \cos \phi_0 \sin \theta_0, \sin\phi_0 \sin \theta_0, \cos \theta_0 \right).
\end{equation}
The instantaneous world-line tangent of the observer fixed to the NS surface at
the point of impact is given in cartesian coordinates by -- IT IS THE AZIMUTHAL VELOCITY? 
\begin{equation}
    \label{eq:tangent_vel}
    \vec{t}_0 = \frac{2 \pi r_{NS}}{P} \frac{\partial \chi^t}{\partial \lambda} \vec{s} \times \vec{r}_0. 
\end{equation}

The cross product is 
\begin{align*}
    % \begin{equation}
        \vec{s} \times \vec{r}_0 = \left(
        -\sin \iota \sin \omega \cos \theta_0 - \sin \iota \cos \omega \sin\phi_0 \sin \theta_0, \right. \\
        \left. \cos \iota \cos \theta_0 - \sin \iota \cos \omega \cos \phi_0 \sin \theta_0, \right. \\
        \left. \cos \iota \sin\phi_0 \sin \theta_0 + \sin \iota \sin \omega \cos \phi_0 \sin \theta_0 \right)
    % \end{equation}  
\end{align*}

Plugging in the cross product into the equation \ref{eq:tangent_vel} and solving for $r-, \theta-, \phi-$ components of the vector, then, in the observer spherical coordinate system the components of the tangent vector
are given by 
\begin{eqnarray}
    \frac{\mathrm{d}\chi^t}{\mathrm{d} \lambda} & \simeq & 1 + \frac{1}{2}  \vRot^2 \sin^2 \theta_0 + \frac{1 + \frac{3}{2} \vRot^2 \sin^2 \theta_0}{\rNS} \\
    \frac{\mathrm{d} \chi^r}{\mathrm{d} \lambda} & = & 0 \\
    \frac{\mathrm{d} \chi^\theta}{\mathrm{d} \lambda} & = & \frac{\mathrm{d} \chi^t}{\mathrm{d}\lambda} \frac{2\pi}{P} \left( - \cos\iota \sin \phi_0 - \cos\phi_0 \sin \iota \sin \omega \right) \\
    \frac{\mathrm{d} \chi^\phi}{\mathrm{d} \lambda} & = & \frac{\mathrm{d} \chi^t}{\mathrm{d}\lambda} \frac{2\pi}{P} \left( -\cos\iota \cos \phi_0 \cot \theta_0  + \sin \iota \left( \cos \omega + \cot \theta_0 \sin \phi_0 \sin \omega \right) \right)
\end{eqnarray}
